Analysis of the implications of the ``Moral Machine'' project as an implementation of the concept of ``coherent, extrapolated volition'' for building clustered trust in autonomous machines

### 2023-02-02 ART_Sekrst/Sekrst.tex
Soloducha article added

\section{Machine-learning basics}
\textit{Machine learning}, as a~part of artificial intelligence and computer science, is a~field of approaches and methods that use data in order to improve their performance on some problems. Artificial intelligence can be seen as a~certain type of philosophical engineering
%\label{ref:RNDwNayypCdyS}(Skansi, 2018, p.vii):
\parencite[][p.vii]{skansi_introduction_2018}: %
 we want ``to build machines that can think, […] understand the meaning, act rationally, cope with uncertainty, [...] handle and talk about objects''. In a~nutshell, we are replicating standard philosophical concepts. It is no wonder that philosophical concepts are deeply embedded in their methods as well but may seem hidden underneath technical layers.

@@

To summarize, a~supervised machine-learning algorithm receives a~set of training data points (a point in space where the axes are the properties given) and labels (row vectors), and in this phase, the algorithm creates a~hyperplane—a decision boundary that helps classify its data points—by adjusting its internal parameters
%\label{ref:RNDwu6xGZUkXK}(Skansi, 2018, pp.55–56).
\parencite[][pp.55–56]{skansi_introduction_2018}. %
 This phase is the training phase that receives inputs as row vectors with corresponding labels (called training samples). In the next, predicting phase, the trained algorithm takes a~number of row vectors, but this time without labels and creates the labels with the hyperplane 
%\label{ref:RNDCI401gVU1c}(Skansi, 2018, p.56).
\parencite[][p.56]{skansi_introduction_2018}. %
 In other words, ``the learner receives a~set of labeled examples as training data and makes predictions for all unseen points, [… a~scenario commonly] associated with classification, regression, and ranking [i.e. ordering items to some criterion] problems'' 
%\label{ref:RND4Xf1QrSB3d}(Mohri, Rostamizadeh and Talwalkar, 2018, p.6).
\parencite[][p.6]{mohri_foundations_2018}.%

Another type of machine learning, \textit{unsupervised learning}, handles various datasets without any explicit instructions or labels. That is, the learner receives unlabeled training data and makes predictions for all unseen points, and ``since, in general, no labeled example is available [...], it can be difficult to quantitatively evaluate the performance of a~learner''
%\label{ref:RNDCBD4vgLA3T}(Mohri, Rostamizadeh and Talwalkar, 2018, p.7).
\parencite[][p.7]{mohri_foundations_2018}. %
 Unsupervised learning encompasses a~broad definition of learning without labels or targets, but this broad definition begs the cognitive question of how we learn without feedback 
%\label{ref:RNDUR3el9pJJt}(Skansi, 2018, p.70).
\parencite[][p.70]{skansi_introduction_2018}. %
 In the previous case, in order for the computer to learn what is a~dog, we had to correctly label dogs or provide a~list of properties in other supervised examples. Here, a~computer is seemingly on its own: for example, neural networks\footnote{See Skansi 
%\label{ref:RNDys19TmMXkr}(2018)
\parencite*[][]{skansi_introduction_2018} %
 for an introduction to deep learning.} tend to automatically find structures in the data by analyzing useful features. Data is often grouped into \textit{clusters}, and then it is easy to see the outliers, anomalies (for example, for fraud detection), associations (for recommender systems), and similar connections.

>>>>

\section{Machine-learning basics}
\textit{Machine learning}, as a~part of artificial intelligence and computer science, is a~field of approaches and methods that use data in order to improve their performance on some problems. Artificial intelligence can be seen as a~certain type of philosophical engineering
%\label{ref:RNDwNayypCdyS}(Skansi, 2018, p.vii):
\parencite[][p.vii]{skansi_introduction_2018}: %
 we want ``to build machines that can think, [\ldots] understand the meaning, act rationally, cope with uncertainty, [\ldots] handle and talk about objects''. In a~nutshell, we are replicating standard philosophical concepts. It is no wonder that philosophical concepts are deeply embedded in their methods as well but may seem hidden underneath technical layers.

@@

To summarize, a~supervised machine-learning algorithm receives a~set of training data points (a point in space where the axes are the properties given) and labels (row vectors), and in this phase, the algorithm creates a~hyperplane—a decision boundary that helps classify its data points—by adjusting its internal parameters
%\label{ref:RNDwu6xGZUkXK}(Skansi, 2018, pp.55–56).
\parencite[][pp.55–56]{skansi_introduction_2018}. %
 This phase is the training phase that receives inputs as row vectors with corresponding labels (called training samples). In the next, predicting phase, the trained algorithm takes a~number of row vectors, but this time without labels and creates the labels with the hyperplane 
%\label{ref:RNDCI401gVU1c}(Skansi, 2018, p.56).
\parencite[][p.56]{skansi_introduction_2018}. %
 In other words, ``the learner receives a~set of labeled examples as training data and makes predictions for all unseen points, [\ldots a~scenario commonly] associated with classification, regression, and ranking [i.e. ordering items to some criterion] problems'' 
%\label{ref:RND4Xf1QrSB3d}(Mohri, Rostamizadeh and Talwalkar, 2018, p.6).
\parencite[][p.6]{mohri_foundations_2018}.%

Another type of machine learning, \textit{unsupervised learning}, handles various datasets without any explicit instructions or labels. That is, the learner receives unlabeled training data and makes predictions for all unseen points, and ``since, in general, no labeled example is available [\ldots], it can be difficult to quantitatively evaluate the performance of a~learner''
%\label{ref:RNDCBD4vgLA3T}(Mohri, Rostamizadeh and Talwalkar, 2018, p.7).
\parencite[][p.7]{mohri_foundations_2018}. %
 Unsupervised learning encompasses a~broad definition of learning without labels or targets, but this broad definition begs the cognitive question of how we learn without feedback 
%\label{ref:RNDUR3el9pJJt}(Skansi, 2018, p.70).
\parencite[][p.70]{skansi_introduction_2018}. %
 In the previous case, in order for the computer to learn what is a~dog, we had to correctly label dogs or provide a~list of properties in other supervised examples. Here, a~computer is seemingly on its own: for example, neural networks\footnote{See Skansi 
%\label{ref:RNDys19TmMXkr}(2018)
\parencite*[][]{skansi_introduction_2018} %
 for an introduction to deep learning.} tend to automatically find structures in the data by analyzing useful features. Data is often grouped into \textit{clusters}, and then it is easy to see the outliers, anomalies (for example, for fraud detection), associations (for recommender systems), and similar connections.

### 2023-02-02 ART_Soloducha/Soloducha.tex
Soloducha article added

\begin{document}
\title{Analysis of the implications of the ``Moral Machine'' project as an implementation of the concept of ``coherent, extrapolated volition'' for building clustered trust in autonomous machines}
\maketitle

Krzysztof Sołoducha PhD, prof. WAT \href{mailto:krzysztof.soloducha@wat.edu.pl}{krzysztof.soloducha@wat.edu.pl}

\section{Abstract}
\begin{quotation}
In this paper, we focus on the analysis of Eliezer Yudkowsky's concept of ``coherent extrapolated volition'' (CEV) as a~response to the need for a~post-conventional, persuasive morality that meets the criteria of active trust in the sense of Anthony Giddens, which could be used in the case of autonomous machines. Based on the analysis of the results of the Moral Machine project, we formulate some guidelines for transformation of the idea of a~coherent extrapolated volition into the concept of a~coherent, extrapolated and clustered volition. The argumentation used in the paper is intended to show that the idea of CEV transformed into its clustered version can be used to build a~technically and socially efficient decision-making pattern database for autonomous machines.

\end{quotation}
\section{Key words}
\begin{quotation}
ethics of artificial intelligence, ethics of autonomous machines, trust in artificial intelligence, moral machine project, coherent extrapolated volition

\end{quotation}
The problem of the ethics of autonomous machines is a~philosophical issue that has emerged together with the dangers of the development of modern technologies that allow the autonomisation of machine operation
%\label{ref:RNDCNwvoNhD5U}(Pasquale, 2016).
\parencite[][]{pasquale_black_2016}. %
 Of course, the classic of considerations on this issue is Isaac Asimov 
%\label{ref:RND8EDGzYgvVE}(2004)
\parencite*[][]{asimov_runaround_2004} %
 and his reflections on the ethics of robots, but at the moment the discussion on this topic is determined primarily by the dynamic development of unsupervised machine learning methods 
%\label{ref:RNDZsSSj9nxYT}(Gryz, 2021).
\parencite[][]{gryz_sztuczna_2021}. %
 They have gained a~second wind (having been theoretically developed in the 1980s) thanks to the development of the Internet and access to large sets of data on which machines can learn and improve algorithms on their own, using formal rules of statistical reasoning and without any kind of supervision 
%\label{ref:RND2WtqnrZTv3}(Burrell, 2016).
\parencite[][]{burrell_how_2016}.%

@@

In this paper we make direct reference to the argument that inductive reasoning can be treated as a~way of solving the is-ought problem. While this reasoning is fraught with the problem of uncertain inference, it is fundamentally consistent with Hume's inductive approach to, for example, the problem of causality. It is an approach that appeals to the weak rationality argument proposed by Searle
%\label{ref:RNDGkPpfSEcym}(1964)
\parencite*[][]{searle_how_1964} %
 and based on the concept of unreliability of purely logical reasoning about duty from description. From this point of view, one can also speak about an attempt to solve the so-called Jorgensen dilemma 
%\label{ref:RNDevSOxEOkT8}(Jörgensen, 1937)
\parencite[][]{jorgensen_imperatives_1937} %
 based on 3 claims:

{}- logically valid reasoning can be made only on the logical sentences (the ones, that can be true or false),

{}- the norms are not logical sentences,

{}- logical correct reasonings are carried out as practical syllogisms.

Therefore the facticity of the practical syllogism is based on reasoning grounded on weak rationality. And this concept of weak rationality used for moral reasoning was used in the Moral Machine project.

@@

In order to find answers to this question, we carried out a~sample survey on 34 students aged between 18 and 39, who were asked about their preferences in this respect. From a~methodological point of view, this is a~survey carried out in the form of an online survey. Its aim was to examine the basic preferences of possible users of autonomous machines in terms of the expected level of clustering. The survey has no ambition to be a~representative poll. Our objective is rather to identify certain trends in user preferences.

Below you find most interesting results of the survey in terms of clustering of preferences of AV users.

Illustration 1

Source: Author's study

The responses are split almost equally between the preferences of the driver and the preferences of the residents of the area in which the vehicle travels, with a~slight advantage to the driver. The conclusion for manufacturers of autonomous devices is therefore that a~device should have an open architecture that allows it to be adapted to the preferred option, unless legal regulations decide otherwise. However, as the survey results show, imposing solutions in this area may end in reduced trust in the device.

Illustration 2

Source: Author's study

@@

\section*{Acknowledgments }
The author wishes to thank anonymous helpful referees whose insightful comments helped improve the quality of the present paper. Most certainly, if there are still some errors remaining, they are my own responsibility.

\section*{References}
Arrow, K.J., 1973. Some Ordinalist-Utilitarian Notes on Rawls's Theory of Justice. \textit{The Journal of Philosophy}, [online] 70(9), pp.245–263. https://doi.org/10.2307/2025006.

Asimov, I., 2004. Runaround. In: \textit{I, robot}, Bantam spectra book, Bantam hardcover edition. New York: Bantam Books. pp.30–55.

Awad, E., Dsouza, S., Shariff, A., Rahwan, I. and Bonnefon, J.-F., 2020. Universals and variations in moral decisions made in 42 countries by 70,000 participants. \textit{Proceedings of the National Academy of Sciences}, [online] 117(5), pp.2332–2337. https://doi.org/10.1073/pnas.1911517117.

Bigman, Y.E. and Gray, K., 2020. Life and death decisions of autonomous vehicles. \textit{Nature}, [online] 579(7797), pp.E1–E2. https://doi.org/10.1038/s41586-020-1987-4.

Bostrom, N., 2016. \textit{Superintelligence: Paths, Dangers, Strategies}. Oxford: Oxford University Press.

Burrell, J., 2016. How the machine ‘thinks': Understanding opacity in machine learning algorithms. \textit{Big Data \& Society}, [online] 3(1), pp.1–12. https://doi.org/10.1177/2053951715622512.

Czyżowska, D., Niemczyński, A. and Kmieć, E., 1993. Formy rozumowania moralnego Polaków w~świetle danych z~badania metodą Lawrence'a Kohlberga. \textit{Kwartalnik Polskiej Psychologii Rozwojowej}, 2(1), pp.19–38.

Dignum, V., 2022. \textit{Responsible Ai: From Principles to Action}. [online] Michigan Institute for Data Science. Available at: {\textless}https://www.youtube.com/watch?v=LwKDOwwJpL4{\textgreater} [Accessed 11 January 2023].

Edmonds, E., 2017. \textit{Americans Feel Unsafe Sharing the Road with Fully Self-Driving Cars}. [online] AAA Newsroom. Available at: {\textless}https://newsroom.aaa.com/2017/03/americans-feel-unsafe-sharing-road-fully-self-driving-cars/{\textgreater} [Accessed 11 January 2023].

Eschenbach, W.J. von, 2021. Transparency and the Black Box Problem: Why We Do Not Trust AI. \textit{Philosophy \& Technology}, [online] 34(4), pp.1607–1622. https://doi.org/10.1007/s13347-021-00477-0.

Foot, P., 2002. \textit{Virtues and Vices}. [online] Oxford University Press. https://doi.org/10.1093/0199252866.001.0001.

Fukuyama, F., 1995. \textit{Trust: the social virtues and the creation of prosperity}. 1st ed. A~Free Press paperbacks book. New York: Free Press.

Fürnkranz, J. and Hüllermeier, E., 2011. Preference Learning: An Introduction. In: J. Fürnkranz and E. Hüllermeier, eds. \textit{Preference Learning}. [online] Berlin, Heidelberg: Springer. pp.1–17. https://doi.org/10.1007/978-3-642-14125-6\_1.

Gellner, E., 2005. \textit{Words and Things: An Examination of, and an Attack on, Linguistic Philosophy}. 1. publ ed. Routledge classics. London: Routledge.

Giddens, A., 1991. \textit{Modernity and Self-Identity: Self and Society in the Late Modern Age}. Stanford, CA: Stanford University Press.

Górnicka, J., 1980. Rozwój moralny w~koncepcji Lawrence'a Kohlberga. \textit{Człowiek i~Światopogląd}, 6, pp.113–123.

Greene, J.D., 2013. \textit{Moral Tribes: Emotion, Reason, and the Gap Between Us and Them}. London: Atlantic Books.

Gryz, J., 2021. \textit{Sztuczna~Inteligencja:~powstanie,~rozwój,~rokowania}. [online] Copernicus Center. Available at: {\textless}https://www.youtube.com/watch?v=3ZDfVgC897k{\textgreater} [Accessed 11 January 2023].

Harsanyi, J.C., 1975. Can the Maximin Principle Serve as a~Basis for Morality? A~Critique of John Rawls's Theory. \textit{The American Political Science Review}, [online] 69(2), pp.594–606. https://doi.org/10.2307/1959090.

Hofstede, G.H., Hofstede, G.J. and Minkov, M., 2010. \textit{Cultures and Organizations: Software of the Mind: Intercultural Cooperation and Its Importance for Survival}. 3rd ed ed. New York: McGraw-Hill.

Hofstede, G.J., 2011. \textit{Geert Hofstede on Culture}. [online] Viewdutch. Available at: {\textless}https://www.youtube.com/watch?v=wdh40kgyYOY{\textgreater} [Accessed 11 January 2023].

Holstein, T. and Dodig-Crnkovic, G., 2018. Avoiding the intrinsic unfairness of the trolley problem. In: \textit{Proceedings of the International Workshop on Software Fairness}. [online] ICSE '18: 40th International Conference on Software Engineering. Gothenburg Sweden: ACM. pp.32–37. https://doi.org/10.1145/3194770.3194772.

Inglehart, R. and Welzel, C., 2005. \textit{Modernization, Cultural Change, and Democracy: The Human Development Sequence}. [online] Cambridge: Cambridge University Press. https://doi.org/10.1017/CBO9780511790881.

Jörgensen, J., 1937. Imperatives and logic. \textit{Erkenntnis}, [online] 7(1), pp.288–296. https://doi.org/10.1007/BF00666538.

Karpus, J., Krüger, A., Verba, J.T., Bahrami, B. and Deroy, O., 2021. Algorithm exploitation: Humans are keen to exploit benevolent AI. \textit{iScience}, [online] 24(6), p.102679. https://doi.org/10.1016/j.isci.2021.102679.

Kohlberg, L., 1958. \textit{The Development of Modes of Moral Thinking and Choice in the Years 10 to 16}. [PhD thesis] University of Chicago. Available at: {\textless}https://www.proquest.com/openview/c503bf59d762abe5818e1b24c484d41a/1?pq-origsite=gscholar\&cbl=18750\&diss=y{\textgreater} [Accessed 11 January 2023].

Lo, T., 2019. \textit{My Amazon Alexa Went Rogue and Ordered Me to Stab Myself in the Heart}. [online] Mirror Online. Available at: {\textless}https://www.mirror.co.uk/news/uk-news/my-amazon-echo-went-rogue-21127994{\textgreater} [Accessed 11 January 2023].

Maroto-Gómez, M., Castro-González, Á., Castillo, J.C., Malfaz, M. and Salichs, M.Á., 2022. An adaptive decision-making system supported on user preference predictions for human–robot interactive communication. \textit{User Modeling and User-Adapted Interaction}, [online] pp.1–45. https://doi.org/10.1007/s11257-022-09321-2.

Miłaszewicz, D., 2016. Zaufanie jako wartość społeczna. \textit{Studia Ekonomiczne}, [online] (259), pp.80–88. Available at: {\textless}http://cejsh.icm.edu.pl/cejsh/element/bwmeta1.element.cejsh-d64b921a-5db7-4208-9eb7-73baaa05f7e4{\textgreater} [Accessed 11 January 2023].

Mirnig, A.G. and Meschtscherjakov, A., 2019. Trolled by the Trolley Problem: On What Matters for Ethical Decision Making in Automated Vehicles. In: \textit{Proceedings of the 2019 CHI Conference on Human Factors in Computing Systems}. [online] CHI '19: CHI Conference on Human Factors in Computing Systems. Glasgow Scotland Uk: ACM. pp.1–10. https://doi.org/10.1145/3290605.3300739.

Nagel, T., 1986. \textit{The View from Nowhere}. 1st ed. Oxford: Oxford University Press.

Nozick, R., 2013. \textit{Anarchy, State, and Utopia}. New York: Basic Books.

Pasquale, F., 2016. \textit{The Black Box Society: The Secret Algorithms That Control Money and Information}. Cambridge, MA: Harvard University Press.

Rawls, J., 1971. \textit{A~Theory of Justice}. [online] Cambridge, MA: The Belknap Press of Harvard University Press. Available at: {\textless}http://www.gbv.de/dms/bowker/toc/9780674880146.pdf{\textgreater} [Accessed 11 January 2023].

Rosenbloom, P., 2015. \textit{On Computing: The Fourth Great Scientific Domain}. Cambridge, MA: The MIT Press.

Searle, J.R., 1964. How to Derive ‘Ought' From ‘Is'. \textit{The Philosophical Review}, [online] 73(1), pp.43–58. https://doi.org/10.2307/2183201.

Wysocki, I., 2021. The problem of indifference and homogeneity in Austrian economics: Nozick's challenge revisited. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (71), pp.9–44. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/554{\textgreater} [Accessed 24 January 2022].

Yudkowsky, E., 2004. \textit{Coherent Extrapolated Volition}. San Francisco, CA: San Francisco.
\end{document}

>>>>

\begin{artengenv}{Krzysztof Sołoducha}
	{Analysis of the implications of the Moral Machine project as an implementation of the concept of coherent extrapolated volition for building clustered trust in autonomous machines}
	{Analysis of the implications of the Moral Machine project\ldots}
	{Analysis of the implications of the Moral Machine project as an implementation of the concept of coherent extrapolated volition for building clustered trust in autonomous machines}
	{Military University of Technology}
	{In this paper, we focus on the analysis of Eliezer Yudkowsky's concept of ``coherent extrapolated volition'' (CEV) as a~response to the need for a~post-conventional, persuasive morality that meets the criteria of active trust in the sense of Anthony Giddens, which could be used in the case of autonomous machines. Based on the analysis of the results of the Moral Machine project, we formulate some guidelines for transformation of the idea of a~coherent extrapolated volition into the concept of a~coherent, extrapolated and clustered volition. The argumentation used in the paper is intended to show that the idea of CEV transformed into its clustered version can be used to build a~technically and socially efficient decision-making pattern database for autonomous machines.
	}
	{ethics of artificial intelligence, ethics of autonomous machines, trust in artificial intelligence, moral machine project, coherent extrapolated volition.}

%\begin{document}
%\title{Analysis of the implications of the ``Moral Machine'' project as an implementation of the concept of ``coherent, extrapolated volition'' for building clustered trust in autonomous machines}
%\maketitle
%
%Krzysztof Sołoducha PhD, prof. WAT \href{mailto:krzysztof.soloducha@wat.edu.pl}{krzysztof.soloducha@wat.edu.pl}
%
%\section{Abstract}
%\begin{quotation}
%In this paper, we focus on the analysis of Eliezer Yudkowsky's concept of ``coherent extrapolated volition'' (CEV) as a~response to the need for a~post-conventional, persuasive morality that meets the criteria of active trust in the sense of Anthony Giddens, which could be used in the case of autonomous machines. Based on the analysis of the results of the Moral Machine project, we formulate some guidelines for transformation of the idea of a~coherent extrapolated volition into the concept of a~coherent, extrapolated and clustered volition. The argumentation used in the paper is intended to show that the idea of CEV transformed into its clustered version can be used to build a~technically and socially efficient decision-making pattern database for autonomous machines.
%
%\end{quotation}
%\section{Key words}
%\begin{quotation}
%ethics of artificial intelligence, ethics of autonomous machines, trust in artificial intelligence, moral machine project, coherent extrapolated volition
%
%\end{quotation}

The problem of the ethics of autonomous machines is a~philosophical issue that has emerged together with the dangers of the development of modern technologies that allow the autonomisation of machine operation
%\label{ref:RNDCNwvoNhD5U}(Pasquale, 2016).
\parencite[][]{pasquale_black_2016}. %
 Of course, the classic of considerations on this issue is Isaac Asimov 
%\label{ref:RND8EDGzYgvVE}(2004)
\parencite*[][]{asimov_runaround_2004} %
 and his reflections on the ethics of robots, but at the moment the discussion on this topic is determined primarily by the dynamic development of unsupervised machine learning methods 
%\label{ref:RNDZsSSj9nxYT}(Gryz, 2021).
\parencite[][]{gryz_sztuczna_2021}. %
 They have gained a~second wind (having been theoretically developed in the 1980s) thanks to the development of the Internet and access to large sets of data on which machines can learn and improve algorithms on their own, using formal rules of statistical reasoning and without any kind of supervision 
%\label{ref:RND2WtqnrZTv3}(Burrell, 2016).
\parencite[][]{burrell_how_2016}.%

@@

In this paper we make direct reference to the argument that inductive reasoning can be treated as a~way of solving the is-ought problem. While this reasoning is fraught with the problem of uncertain inference, it is fundamentally consistent with Hume's inductive approach to, for example, the problem of causality. It is an approach that appeals to the weak rationality argument proposed by Searle
%\label{ref:RNDGkPpfSEcym}(1964)
\parencite*[][]{searle_how_1964} %
 and based on the concept of unreliability of purely logical reasoning about duty from description. From this point of view, one can also speak about an attempt to solve the so-called Jorgensen dilemma 
%\label{ref:RNDevSOxEOkT8}(Jörgensen, 1937)
\parencite[][]{jorgensen_imperatives_1937} %
 based on 3 claims:
\begin{itemize}
%
%{}- logically valid reasoning can be made only on the logical sentences (the ones, that can be true or false),
%
%{}- the norms are not logical sentences,
%
%{}- logical correct reasonings are carried out as practical syllogisms.
\item logically valid reasoning can be made only on the logical sentences (the ones, that can be true or false),
\item the norms are not logical sentences,
\item logical correct reasonings are carried out as practical syllogisms.
\end{itemize}
Therefore the facticity of the practical syllogism is based on reasoning grounded on weak rationality. And this concept of weak rationality used for moral reasoning was used in the Moral Machine project.

@@

In order to find answers to this question, we carried out a~sample survey on 34 students aged between 18 and 39, who were asked about their preferences in this respect. From a~methodological point of view, this is a~survey carried out in the form of an online survey. Its aim was to examine the basic preferences of possible users of autonomous machines in terms of the expected level of clustering. The survey has no ambition to be a~representative poll. Our objective is rather to identify certain trends in user preferences.
%Below you find most interesting results of the survey in terms of clustering of preferences of AV users.
\enlargethispage{2\baselineskip}
\begin{figure}[H]
 \begin{center}
 \includegraphics[width=.8\textwidth]{ART_Soloducha/illustration1.pdf}%
 \end{center}%
 \caption{Most interesting results of the survey in terms of clustering of preferences of AV users. Source: Author's study.}\label{sol-ill1}
\end{figure}
%Illustration 1

The responses are split almost equally between the preferences of the driver and the preferences of the residents of the area in which the vehicle travels, with a~slight advantage to the driver. The conclusion for manufacturers of autonomous devices is therefore that a~device should have an open architecture that allows it to be adapted to the preferred option, unless legal regulations decide otherwise. However, as the survey results show, imposing solutions in this area may end in reduced trust in the device.
\begin{figure}
 \begin{center}
 \includegraphics[width=.8\textwidth]{ART_Soloducha/illustration2.pdf}%
 \end{center}%
 \caption{Level of clustering survey. Source: Author's study.}\label{sol-ill2}
\end{figure}
%Illustration 2
%Source: Author's study

@@

\paragraph{Acknowledgments.}
The author wishes to thank anonymous helpful referees whose insightful comments helped improve the quality of the present paper. Most certainly, if there are still some errors remaining, they are my own responsibility.

\end{artengenv}

%\section*{References}
%Arrow, K.J., 1973. Some Ordinalist-Utilitarian Notes on Rawls's Theory of Justice. \textit{The Journal of Philosophy}, [online] 70(9), pp.245–263. https://doi.org/10.2307/2025006.
%
%Asimov, I., 2004. Runaround. In: \textit{I, robot}, Bantam spectra book, Bantam hardcover edition. New York: Bantam Books. pp.30–55.
%
%Awad, E., Dsouza, S., Shariff, A., Rahwan, I. and Bonnefon, J.-F., 2020. Universals and variations in moral decisions made in 42 countries by 70,000 participants. \textit{Proceedings of the National Academy of Sciences}, [online] 117(5), pp.2332–2337. https://doi.org/10.1073/pnas.1911517117.
%
%Bigman, Y.E. and Gray, K., 2020. Life and death decisions of autonomous vehicles. \textit{Nature}, [online] 579(7797), pp.E1–E2. https://doi.org/10.1038/s41586-020-1987-4.
%
%Bostrom, N., 2016. \textit{Superintelligence: Paths, Dangers, Strategies}. Oxford: Oxford University Press.
%
%Burrell, J., 2016. How the machine ‘thinks': Understanding opacity in machine learning algorithms. \textit{Big Data \& Society}, [online] 3(1), pp.1–12. https://doi.org/10.1177/2053951715622512.
%
%Czyżowska, D., Niemczyński, A. and Kmieć, E., 1993. Formy rozumowania moralnego Polaków w~świetle danych z~badania metodą Lawrence'a Kohlberga. \textit{Kwartalnik Polskiej Psychologii Rozwojowej}, 2(1), pp.19–38.
%
%Dignum, V., 2022. \textit{Responsible Ai: From Principles to Action}. [online] Michigan Institute for Data Science. Available at: {\textless}https://www.youtube.com/watch?v=LwKDOwwJpL4{\textgreater} [Accessed 11 January 2023].
%
%Edmonds, E., 2017. \textit{Americans Feel Unsafe Sharing the Road with Fully Self-Driving Cars}. [online] AAA Newsroom. Available at: {\textless}https://newsroom.aaa.com/2017/03/americans-feel-unsafe-sharing-road-fully-self-driving-cars/{\textgreater} [Accessed 11 January 2023].
%
%Eschenbach, W.J. von, 2021. Transparency and the Black Box Problem: Why We Do Not Trust AI. \textit{Philosophy \& Technology}, [online] 34(4), pp.1607–1622. https://doi.org/10.1007/s13347-021-00477-0.
%
%Foot, P., 2002. \textit{Virtues and Vices}. [online] Oxford University Press. https://doi.org/10.1093/0199252866.001.0001.
%
%Fukuyama, F., 1995. \textit{Trust: the social virtues and the creation of prosperity}. 1st ed. A~Free Press paperbacks book. New York: Free Press.
%
%Fürnkranz, J. and Hüllermeier, E., 2011. Preference Learning: An Introduction. In: J. Fürnkranz and E. Hüllermeier, eds. \textit{Preference Learning}. [online] Berlin, Heidelberg: Springer. pp.1–17. https://doi.org/10.1007/978-3-642-14125-6\_1.
%
%Gellner, E., 2005. \textit{Words and Things: An Examination of, and an Attack on, Linguistic Philosophy}. 1. publ ed. Routledge classics. London: Routledge.
%
%Giddens, A., 1991. \textit{Modernity and Self-Identity: Self and Society in the Late Modern Age}. Stanford, CA: Stanford University Press.
%
%Górnicka, J., 1980. Rozwój moralny w~koncepcji Lawrence'a Kohlberga. \textit{Człowiek i~Światopogląd}, 6, pp.113–123.
%
%Greene, J.D., 2013. \textit{Moral Tribes: Emotion, Reason, and the Gap Between Us and Them}. London: Atlantic Books.
%
%Gryz, J., 2021. \textit{Sztuczna~Inteligencja:~powstanie,~rozwój,~rokowania}. [online] Copernicus Center. Available at: {\textless}https://www.youtube.com/watch?v=3ZDfVgC897k{\textgreater} [Accessed 11 January 2023].
%
%Harsanyi, J.C., 1975. Can the Maximin Principle Serve as a~Basis for Morality? A~Critique of John Rawls's Theory. \textit{The American Political Science Review}, [online] 69(2), pp.594–606. https://doi.org/10.2307/1959090.
%
%Hofstede, G.H., Hofstede, G.J. and Minkov, M., 2010. \textit{Cultures and Organizations: Software of the Mind: Intercultural Cooperation and Its Importance for Survival}. 3rd ed ed. New York: McGraw-Hill.
%
%Hofstede, G.J., 2011. \textit{Geert Hofstede on Culture}. [online] Viewdutch. Available at: {\textless}https://www.youtube.com/watch?v=wdh40kgyYOY{\textgreater} [Accessed 11 January 2023].
%
%Holstein, T. and Dodig-Crnkovic, G., 2018. Avoiding the intrinsic unfairness of the trolley problem. In: \textit{Proceedings of the International Workshop on Software Fairness}. [online] ICSE '18: 40th International Conference on Software Engineering. Gothenburg Sweden: ACM. pp.32–37. https://doi.org/10.1145/3194770.3194772.
%
%Inglehart, R. and Welzel, C., 2005. \textit{Modernization, Cultural Change, and Democracy: The Human Development Sequence}. [online] Cambridge: Cambridge University Press. https://doi.org/10.1017/CBO9780511790881.
%
%Jörgensen, J., 1937. Imperatives and logic. \textit{Erkenntnis}, [online] 7(1), pp.288–296. https://doi.org/10.1007/BF00666538.
%
%Karpus, J., Krüger, A., Verba, J.T., Bahrami, B. and Deroy, O., 2021. Algorithm exploitation: Humans are keen to exploit benevolent AI. \textit{iScience}, [online] 24(6), p.102679. https://doi.org/10.1016/j.isci.2021.102679.
%
%Kohlberg, L., 1958. \textit{The Development of Modes of Moral Thinking and Choice in the Years 10 to 16}. [PhD thesis] University of Chicago. Available at: {\textless}https://www.proquest.com/openview/c503bf59d762abe5818e1b24c484d41a/1?pq-origsite=gscholar\&cbl=18750\&diss=y{\textgreater} [Accessed 11 January 2023].
%
%Lo, T., 2019. \textit{My Amazon Alexa Went Rogue and Ordered Me to Stab Myself in the Heart}. [online] Mirror Online. Available at: {\textless}https://www.mirror.co.uk/news/uk-news/my-amazon-echo-went-rogue-21127994{\textgreater} [Accessed 11 January 2023].
%
%Maroto-Gómez, M., Castro-González, Á., Castillo, J.C., Malfaz, M. and Salichs, M.Á., 2022. An adaptive decision-making system supported on user preference predictions for human–robot interactive communication. \textit{User Modeling and User-Adapted Interaction}, [online] pp.1–45. https://doi.org/10.1007/s11257-022-09321-2.
%
%Miłaszewicz, D., 2016. Zaufanie jako wartość społeczna. \textit{Studia Ekonomiczne}, [online] (259), pp.80–88. Available at: {\textless}http://cejsh.icm.edu.pl/cejsh/element/bwmeta1.element.cejsh-d64b921a-5db7-4208-9eb7-73baaa05f7e4{\textgreater} [Accessed 11 January 2023].
%
%Mirnig, A.G. and Meschtscherjakov, A., 2019. Trolled by the Trolley Problem: On What Matters for Ethical Decision Making in Automated Vehicles. In: \textit{Proceedings of the 2019 CHI Conference on Human Factors in Computing Systems}. [online] CHI '19: CHI Conference on Human Factors in Computing Systems. Glasgow Scotland Uk: ACM. pp.1–10. https://doi.org/10.1145/3290605.3300739.
%
%Nagel, T., 1986. \textit{The View from Nowhere}. 1st ed. Oxford: Oxford University Press.
%
%Nozick, R., 2013. \textit{Anarchy, State, and Utopia}. New York: Basic Books.
%
%Pasquale, F., 2016. \textit{The Black Box Society: The Secret Algorithms That Control Money and Information}. Cambridge, MA: Harvard University Press.
%
%Rawls, J., 1971. \textit{A~Theory of Justice}. [online] Cambridge, MA: The Belknap Press of Harvard University Press. Available at: {\textless}http://www.gbv.de/dms/bowker/toc/9780674880146.pdf{\textgreater} [Accessed 11 January 2023].
%
%Rosenbloom, P., 2015. \textit{On Computing: The Fourth Great Scientific Domain}. Cambridge, MA: The MIT Press.
%
%Searle, J.R., 1964. How to Derive ‘Ought' From ‘Is'. \textit{The Philosophical Review}, [online] 73(1), pp.43–58. https://doi.org/10.2307/2183201.
%
%Wysocki, I., 2021. The problem of indifference and homogeneity in Austrian economics: Nozick's challenge revisited. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (71), pp.9–44. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/554{\textgreater} [Accessed 24 January 2022].
%
%Yudkowsky, E., 2004. \textit{Coherent Extrapolated Volition}. San Francisco, CA: San Francisco.
%\end{document}
